Explicit Operations on Parametrised Polytopes

### 2017-02-01 jems-draft.tex
minor edits
Edited the abstract, introduction, conclusions and references.
Also changed (5.5)-(5.10) so that they are equivalent to eqns in Sect.
5.2 of thesis, and changed the end of Section 5.
Changed captions of Fig 1 and 2.

\begin{abstract}
The Pontryagin set difference is an operation between two sets that is used to calculate sets of safe operation in diverse applications including game theory and robust control.
%
We consider a particular class of set-valued maps; for this we introduce the property of parametric convexity and present some of its analytic framework.
%
In order to guarantee the convexity of the difference of a convex set and a set-valued map it proves necessary and sufficient that the set-valued map is parametrically convex.
%
For generic piecewise polyhedral set-valued maps we show that parametric convexity is equivalent to a constant combinatorial structure of all its realisations.
%
We present a computational method for verifying parametric convexity of a piecewise polyhedral set-valued map as well as a way to calculate the parametric Pontryagin difference between a polyhedron and a piecewise polyhedral set-valued map.
%
Two examples are used to illustrate the presented methodology.

@@

\section{Computational Methods}\label{sec:computational:methods}
%
%
%
\subsection{Determination of Parametric Convexity}
In the previous sections we have derived that for the Pontryagin difference between any convex set $S$ and a set-valued map~$\mathcal W$ to be convex itself we require the parametric convexity property for~$\mathcal W$.
%
We have also seen that for set-valued maps with polyhedral realisations $\mathcal W(p)$, the number of vertices and rays must not change throughout the parameter space~$Y$ in order that $\mathcal W(p)$ is parametrically convex.
%
This might seem like a computationally challenging endeavour, however we will see that it suffices to solve a finite number of calculations.
%
For notational convenience let $\phi_i(p) = \max_k\{b_{i,k}+c_{i,k}p\}$ denote the right hand side of~\eqref{eq:definition:PWA:polytopic:set:general}, i.e. $\mathcal W(p) = \{z\in\mathbb R^n: a_i z\leq \phi_i(p)\;\forall i\in\mathbb N_m\}$.
%
It is equivalent to study~$\phi_i$ and its \emph{epigraph} 
%
$$
	\epi(\phi_i) = \{(p,t)\in Y\times\mathbb R: \phi_i(p)\leq t\}
$$
%
see e.g.~\cite{Gorokhovik:1993}.
%
On each $d$-dimensional face 
%
\begin{equation}\label{eq:conv:hull:facet:of:epigraph}
F_j = \underset{f}{\conv}\left\{\left(\begin{array}{c}p_{j,f}\\ t_{j,f}\end{array}\right)\right\}
\end{equation}
%
of $\epi(\phi_i)$ there exists a maximiser~$k_j$ such that~$\phi_i$ is defined by~$\phi_i(p)=b_{i,k_j}+c_{i,k_j}p$ for all $p\in\pi_d(F_j)$ or equivalently~$\phi_i(p)=\sum_{f}\lambda_f t_{j,f}$ with ${\bf{0}}\leq\lambda\leq{\bf{1}}$ such that $p=\sum_{f}\lambda_f p_{j,f}$.
%
Notice that if~\eqref{eq:conv:hull:facet:of:epigraph} is redundant, i.e. not all elements are vertices of $F_j$, then the statement is still true, and we can write $\phi_i(p)$ as a convex combination of points.

@@

$$\abs{\text{vert}(\mathcal W(p))} = \text{const.}\quad \forall p\in\mathcal D$$.

@@

then Corollary~\ref{thm:p:convexity:PWA:set:constant:num:verts} holds and $\mathcal W(p)$ is parametrically convex, furthermore enumerating the vertices of $\mathcal W(p)$ for any fixed $p\in Y$ yields all vertex defining index sets $\mathcal A_\kappa\subset\mathbb N_m$ which remain unchanged for all $p\neq\tilde p\in Y$ according to Corollary~\ref{thm:combinatorical:equivalence:alternative}.

@@

The constancy of the active inequalities defining vertices of~$\mathcal W(p)$ is particularly useful when we compute the parametric Pontryagin difference $S\ominus\mathcal W(S)$:

@@

For the set $S=\{p\in\mathbb R^d:\Lambda_i p\leq\lambda_i,i\in\mathbb N_q\}$ each facet $P_i=\{p:\Lambda_i p=\lambda_i\wedge\Lambda_j p\leq\lambda_j,j\neq i\}$ defines facets of the set difference.

@@

The term~$(\dagger)$ is a mpLP, we are less interested in the objective value of~$(\dagger)$ but rather in the shape of its solution over~$p$ since it defines the shape of the set of points 
which can only just be taken to $P_i$ for an extremal value of $w\in\mathcal W(p)$ and therefore form the boundary of $S\ominus \mathcal W(S)$.

@@

  & B_i+C_ip\leq t_i\\
  & j\in\mathbb N_N

@@

further, the solution to~\eqref{eq:rewriting:pontryagin:diff:without:facet:constraint} is only relevant for such~$p$ that can be taken to~$P_i$ for some $i\in\mathbb N_q$.

@@

We can add that constraint to~\eqref{eq:rewriting:pontryagin:diff:without:facet:constraint} in order to obtain a linear program that is only feasible on the points which can be taken to $P_i$ for feasible $w$:

@@

  & B_j+C_j p\leq t_j\\

@@

  &l\neq i, j\in\mathbb N_N

@@

   k\in\mathbb N_q

@@

but this is the definition of the projection onto $\mathbb R^d$.

@@

With this the boundary of $S\ominus \mathcal W(S)$ is given as the union of projections:

@@

  \text{vert}(S\ominus\mathcal W(S)) = \bigcup_{i\in\{1,\dots,q\}}\pi_d(\text{vert}(\tilde P_i))

@@

We further propose an alternative way of computing $S\ominus\mathcal W(S)$:
%
Instead of using~$\tilde P_i$ which are points~$p$ that can be taken to $P_i$ for a feasible $w\in\mathcal W(p)$ and the values of the~$\phi_i(p)$ and so on, we obtain all feasible~$p$ such that a feasible $w\in\mathcal W(p)$ exists such that $p+w$ lies on the boundary of $S$.
%
\begin{equation}
  L = \left\{(p,t,\tau):\begin{array}{rcl}
  \Lambda_iT_jt&\leq&\tau_i\\
  B_i+C_i p&\leq& t_i\\
  \Lambda_i p + \tau_i&\leq&\lambda_i\\
  \end{array},\begin{array}{rcl}
  j&\in&\mathbb N_N\\
  i&\in&\mathbb N_q\end{array}
  \right\}
\end{equation}
%
The set~$L$ contains all such $p$ for which there exists a feasible $w\in\mathcal W(p)$ such that $p+w\in S$ and hence $\pi_d(L)=S\ominus\mathcal W(S)$, which does not require a vertex enumeration.
%
%
%
%
%
\section{Numerical Examples}\label{sec:numerical:examples}
\subsection{Norm bound sets}\label{ssec:example:one}
%
Consider the set~$X = \{x\in\mathbb R^2: \abs{x_1}\leq 1\wedge \abs{x_2}\leq 1\}$ and the set-valued map $\mathcal W(x) = \{w\in\mathbb R^2: \norm{w}_P\leq\kappa\norm{x}_P\}$.
%
Where $\norm{\cdot}_P$ is a polytopic approximation to the euclidean norm~$\norm{\cdot}_2$, i.e.
%
$$
c_1 \norm{x}_P\leq\norm{x}_2\leq c_2\norm{x}_P.
$$
%
For example
%
$$\begin{aligned}
\norm{x}_P &= \max_{k\leq n}\left\{\sin\left(\frac{2\pi}{n-1}k\right) x_1 + \cos\left(\frac{2\pi}{n-1}k\right) x_2\right\}\\
 &= \max_{k\leq n}\{F_k x\}.
\end{aligned}$$
%
For this choice we find $c_1 = \cos\left(\frac{\pi}{n-1}\right)$ and $c_2=1$.
%
Here~$\mathcal W(x)$ is a scaled approximation of an euclidean ball and therefore obviously parametrically convex with a fixed combinatorial structure.
%
In fact, we find the~$n$ vertices given by the solution to
%
$$
  \left(\begin{array}{cc}
  \sin\left(\frac{2\pi}{n-1}(k-1)\right) & \cos\left(\frac{2\pi}{n-1}(k-1)\right) \\
  \sin\left(\frac{2\pi}{n-1}k\right) & \cos\left(\frac{2\pi}{n-1}k\right)
  \end{array}
  \right) w_k = {\bf{1}}\norm{x}_P
$$
%
for $k=1,\dots,n$.

@@

\subsection{Element wise bound sets}\label{exp:second:example}
%
For another example we take inspiration from a robust control scenario.
%
We want to approximate the maximal positive invariant set for a system of the type~$x^+=Ax + f(x)$ for $x\in\mathcal X$, i.e. the largest set of points~$\mathcal X^\infty\subseteq\mathcal X$ such that $x\in\mathcal X^\infty\Rightarrow x^+\in\mathcal X^\infty$, see e.g.~\cite{blanchini:2007}.
%
For general nonlinear systems~$x^+=Ax + f(x)$ the maximal positive invariant set cannot be calculated, but assume we find piecewise affine bounds on the nonlinearity, i.e. $\min_{k\leq M_i}\{h_{i,k}^x x + h_{i,k}^c\}\leq f_i(x)\leq\max_{k\leq N_i}\{H_{i,k}^x x + H_{i,k}^c\}$ for all $x\in\mathcal X$ and all elements of~$f(x)$.
%
We can now study the system $x^+=Ax+w$ with $w\in\mathcal W(x) = \{w:\min_{k\leq M_i}\{h_{i,k}^x x + h_{i,k}^c\}\leq w_i\leq\max_{k\leq N_i}\{H_{i,k}^x x + H_{i,k}^c\}\,\forall i\in\mathbb N_d\}$.
%
To bound the maximal positive invariant set for the nonlinear system we compute the maximal robust positive invariant set for the perturbed linear system $x^+=Ax+w$, i.e. the set of states~$\mathscr X^\infty\subseteq\mathcal X$ such that $x\in\mathscr X^\infty$ implies $x^+\in\mathscr X^\infty$ for all possible $w\in\mathcal W(x)$.
%
The algorithm to determine the maximal robust positive invariant sets recursively intersects the current set iterate~$X_l$ with the set of all possible successor states starting from the current iterate $X_{l+1}=X_l\cap\left\{\bigcup_{x\in X_l,w\in\mathcal W(x)}\{Ax+w\}\right\}$.
%
We will only present a single step of this procedure.
%
Notice that for a constant set-valued map~$\mathcal W(x) = \mathcal W$ the iteration can be summarised as $X_{l+1} = X_l\cap A^{-1}(X_l\ominus\mathcal W)$, for a set valued map a similar compact notation also holds, however it is more confusing than helpful:~$X_{l+1}=X_l\cap(A^{-1}X_l\ominus A^{-1}\mathcal W(A^{-1}X_l))$.

@@

\begin{figure}
\centering
\begin{tikzpicture}[scale=1.8]
\draw[-latex'] (-2.1,0) -- (2.1,0) node[above] {$t$};
\draw[-latex'] (0,-.9) -- (0,.9) node[right,blue] {$f_1(t)$};
\foreach \x in {-2,-1.5556,...,2} \draw (\x,0.05) -- (\x,-.05);
\draw (-2,-.05) node[below] {$-2$};
\draw (2,-.05) node[below] {$2$};
\foreach \y in {-.8,-.4,...,.8} \draw (0.05,\y) -- (-.05,\y);
\draw (.05,-.8) node[right] {$-0.8$};
\draw (.05,-.4) node[right] {$-0.4$};
\draw[scale=1,domain=-2:2,smooth,variable=\x,blue] plot ({\x},{\x*\x*\x/10});
\draw (-2.0000, -0.8000) -- (-1.5556, -0.3764) -- (-1.1111, -0.1372) -- (-0.6667, -0.0296) -- (-0.2222, -0.0011) -- (0,0); 
\draw (0,0) -- (0.2222, 0.0011) -- (0.6667, 0.0296) -- (1.1111, 0.1372) -- (1.5556, 0.3764) -- (2.0000, 0.8000);
\fill[pattern = north east lines, opacity=.5] (-2.0000, -0.8000) -- (-1.5556, -0.3764) -- (-1.1111, -0.1372) -- (-0.6667, -0.0296) -- (-0.2222, -0.0011) -- (0,0) -- (-2,0) -- cycle; 
\fill[pattern = north east lines, opacity=.5] (0,0) -- (0.2222, 0.0011) -- (0.6667, 0.0296) -- (1.1111, 0.1372) -- (1.5556, 0.3764) -- (2.0000, 0.8000) -- (2,0) -- cycle;
\end{tikzpicture}
\caption{$f_1(t=\frac{x_1}{2}+x_2)$ and its approximation by 9 secants used in Example~\ref{exp:second:example}.}
\label{fig:approximation:f1:p:diff}
\end{figure}

>>>>

\begin{abstract}
The Pontryagin set difference is an operation between two sets that is used to calculate sets of safe operation in diverse applications including game theory and robust control.
%
We consider a particular class of set-valued maps; for this we introduce the property of parametric convexity and present some of its analytic framework.
%
In order to guarantee the convexity of the difference of a convex set and a set-valued map it proves necessary and sufficient that the set-valued map is parametrically convex.
%
For generic piecewise polyhedral set-valued maps we show that parametric convexity is equivalent to a constant combinatorial structure of all its realisations.
%
We present a computational method for verifying parametric convexity of a piecewise polyhedral set-valued map as well as a way to calculate the parametric Pontryagin difference between a polyhedron and a piecewise polyhedral set-valued map.
%
Two examples are given to illustrate the presented methodology.

@@

\section{Computational Methods}\label{sec:computational:methods}
%
%
%
\subsection{Determination of Parametric Convexity}
%In the previous sections it was demonstrated that a set-valued
%map~$\mathcal W$ must be parametrically convex in order that the
%Pontryagin difference between any convex set $S$ and $\mathcal W$ is convex.
%
In Section~3 it was demonstrated that, in order for the
Pontryagin difference between any convex set $S$ and a set-valued
map~$\mathcal W$ to be convex, $\mathcal W$ must be parametrically
convex.
%
We have also seen that for a set-valued map with polyhedral realisations $\mathcal W(p)$, the number of vertices and rays must not change throughout the parameter space~$Y$ in order that $\mathcal W(p)$ is parametrically convex.
%
Checking this condition may seem computationally challenging, however
we show in this section that it suffices to perform a finite number of calculations.
%
For notational convenience let $\phi_i(p) = \max_k\{b_{i,k}+c_{i,k}p\}$ denote the right hand side of~\eqref{eq:definition:PWA:polytopic:set:general}, i.e. $\mathcal W(p) = \{z\in\mathbb R^n: a_i z\leq \phi_i(p)\;\forall i\in\mathbb N_m\}$.
%
It is equivalent to study~$\phi_i$ and its \emph{epigraph} 
%
$$
	\epi(\phi_i) = \{(p,t)\in Y\times\mathbb R: \phi_i(p)\leq t\}
$$
%
see e.g.~\cite{Gorokhovik:1993}.
%
On each $d$-dimensional face 
%
\begin{equation}\label{eq:conv:hull:facet:of:epigraph}
F_j = \underset{f}{\conv}\left\{\left(\begin{array}{c}p_{j,f}\\ t_{j,f}\end{array}\right)\right\}
\end{equation}
%
of $\epi(\phi_i)$ there exists a maximiser~$k_j$ such that~$\phi_i$ is defined by~$\phi_i(p)=b_{i,k_j}+c_{i,k_j}p$ for all $p\in\pi_d(F_j)$ or equivalently~$\phi_i(p)=\sum_{f}\lambda_f t_{j,f}$ with ${\bf{0}}\leq\lambda\leq{\bf{1}}$ such that $p=\sum_{f}\lambda_f p_{j,f}$.
%
Notice that if~\eqref{eq:conv:hull:facet:of:epigraph} is redundant, i.e. not all elements are vertices of $F_j$, then the statement is still true, and we can write $\phi_i(p)$ as a convex combination of points.

@@

$$\abs{\text{vert}(\mathcal W(p))} = c_{\mathcal D} \quad \forall p\in\mathcal D$$

@@

for some constant $c_{\mathcal D}$. If this is the case, then
Corollary~\ref{thm:p:convexity:PWA:set:constant:num:verts} holds and
$\mathcal W(p)$ is parametrically convex. Furthermore enumerating the vertices of $\mathcal W(p)$ for any fixed $p\in Y$ yields all vertices defining index sets $\mathcal A_\kappa\subset\mathbb N_m$ which remain unchanged for all $p\neq\tilde p\in Y$ according to Corollary~\ref{thm:combinatorical:equivalence:alternative}.

@@

The constancy of the active inequalities defining vertices of~$\mathcal W(p)$ is particularly useful when we compute the parametric Pontryagin difference $S\ominus\mathcal W(S)$.

@@

In particular for the set $S=\{p\in\mathbb R^d:\Lambda_i p\leq\lambda_i,i\in\mathbb N_q\}$ each facet $P_i=\{p:\Lambda_i p=\lambda_i\wedge\Lambda_j p\leq\lambda_j,j\neq i\}$ defines facets of the set difference.

@@

The term~$(\dagger)$ is a mpLP, however we are less interested in the
objective value of~$(\dagger)$ than the shape of its solution
over~$p$, since this defines the shape of the set of points 
that can only just be taken to $P_i$ for an extremal value of $w\in\mathcal W(p)$ and hence form the boundary of $S\ominus \mathcal W(S)$.

@@

  & B_k+C_k p \leq t_k\\
  & j\in\mathbb N_N, \ k\in\mathbb N_m 

@@

Furthermore the solution to~\eqref{eq:rewriting:pontryagin:diff:without:facet:constraint} is only relevant for such~$p$ that can be taken to~$P_i$ for some $i\in\mathbb N_q$.

@@

By including this constraint in~\eqref{eq:rewriting:pontryagin:diff:without:facet:constraint}, we obtain a linear program that is only feasible on the points which can be taken to $P_i$ for feasible $w$:

@@

  & B_k+C_k p\leq t_k\\

@@

  &l\neq i, \ j\in\mathbb N_N, \ k\in\mathbb N_m 

@@

   k\in\mathbb N_m

@@

%but this is the definition of the projection onto $\mathbb R^d$.

@@

Therefore the boundary of $S\ominus \mathcal W(S)$ can be expressed as the union of projections:

@@

  \text{vert}(S\ominus\mathcal W(S)) =
  \bigcup_{i\in\{1,\dots,q\}}\pi_d(\text{vert}(\tilde P_i)) .

@@

We conclude this section by proposing an alternative way of computing $S\ominus\mathcal W(S)$.
%
%Instead of using~$\tilde P_i$, which consists of points~$p$ that can be taken
%to $P_i$ for a feasible $w\in\mathcal W(p)$, and the values of
%$\phi_i(p)$, 
%we obtain all feasible~$p$ such that a feasible $w\in\mathcal W(p)$
%exists such that $p+w$ lies on the boundary of $S$.
%
%To directly obtain all feasible~$p$ such that a feasible $w\in\mathcal W(p)$
%exists such that $p+w$ lies on the boundary of $S$, c
Consider the set
%
\begin{equation}
  L = \left\{(p,t,\tau):\begin{array}{rcl}
  \Lambda_iT_jt&\leq&\tau_i\\
  B_k+C_k p&\leq& t_k\\
  \Lambda_i p + \tau_i&\leq&\lambda_i\\
  \end{array},\begin{array}{l}
  j\in\mathbb N_N\\
  k\in\mathbb N_m\\
  i\in\mathbb N_q\end{array}
  \right\} .
\end{equation}
%
The set~$L$ contains all points $p$ for which there exists a feasible $w\in\mathcal W(p)$ such that $p+w\in S$ and hence $\pi_d(L)=S\ominus\mathcal W(S)$, which does not require a vertex enumeration.
%
Notice that although both of the methods described in this section
provide the same result, the first method requires the vertex
enumeration of $q$ $(d + m + 1)$-dimensional polyhedrons, whereas the
latter requires the projection of a $(d + m + q)$-dimensional polyhedron
onto $\mathbb R^d$. Depending on the complexity of $S$, the appropriate method may be chosen.
%
%
%
%
\section{Numerical Examples}\label{sec:numerical:examples}
\subsection{Norm bound sets}\label{ssec:example:one}
%
Consider the set~$X = \{x\in\mathbb R^2: \abs{x_1}\leq 1\wedge \abs{x_2}\leq 1\}$ and the set-valued map $\mathcal W(x) = \{w\in\mathbb R^2: \norm{w}_P\leq\kappa\norm{x}_P\}$.
%
Where $\norm{\cdot}_P$ is a polytopic approximation to the euclidean norm~$\norm{\cdot}_2$, i.e.
%
$$
c_1 \norm{x}_P\leq\norm{x}_2\leq c_2\norm{x}_P.
$$
%
For example
%
$$\begin{aligned}
\norm{x}_P &= \max_{k\leq n}\left\{\sin\left(\frac{2\pi}{n-1}k\right) x_1 + \cos\left(\frac{2\pi}{n-1}k\right) x_2\right\}\\
 &= \max_{k\leq n}\{F_k x\}.
\end{aligned}$$
%
For this choice we find $c_1 = \cos\left(\frac{\pi}{n-1}\right)$ and $c_2=1$.
%
Here~$\mathcal W(x)$ is a scaled approximation of an euclidean ball and therefore obviously parametrically convex with a fixed combinatorial structure.
%
In fact, we find the~$n$ vertices given by the solution to
%
$$
  \left(\begin{array}{cc}
  \sin\left(\frac{2\pi}{n-1}(k-1)\right) & \cos\left(\frac{2\pi}{n-1}(k-1)\right) \\
  \sin\left(\frac{2\pi}{n-1}k\right) & \cos\left(\frac{2\pi}{n-1}k\right)
  \end{array}
  \right) w_k = {\bf{1}}\norm{x}_P
$$
%
for $k=1,\dots,n$.

@@

\subsection{Elementwise bound sets}\label{exp:second:example}
%
For another example we take inspiration from a robust control scenario.
%
We want to approximate the maximal positive invariant set for a system of the type~$x^+=Ax + f(x)$ for $x\in\mathcal X$, i.e. the largest set of points~$\mathcal X^\infty\subseteq\mathcal X$ such that $x\in\mathcal X^\infty\Rightarrow x^+\in\mathcal X^\infty$, see e.g.~\cite{blanchini:2007}.
%
For general nonlinear systems~$x^+=Ax + f(x)$ the maximal positive invariant set cannot be calculated, but assume we find piecewise affine bounds on the nonlinearity, i.e. $\min_{k\leq M_i}\{h_{i,k}^x x + h_{i,k}^c\}\leq f_i(x)\leq\max_{k\leq N_i}\{H_{i,k}^x x + H_{i,k}^c\}$ for all $x\in\mathcal X$ and all elements of~$f(x)$.
%
We can now study the system $x^+=Ax+w$ with $w\in\mathcal W(x) = \{w:\min_{k\leq M_i}\{h_{i,k}^x x + h_{i,k}^c\}\leq w_i\leq\max_{k\leq N_i}\{H_{i,k}^x x + H_{i,k}^c\}\,\forall i\in\mathbb N_d\}$.
%
To bound the maximal positive invariant set for the nonlinear system we compute the maximal robust positive invariant set for the perturbed linear system $x^+=Ax+w$, i.e. the set of states~$\mathscr X^\infty\subseteq\mathcal X$ such that $x\in\mathscr X^\infty$ implies $x^+\in\mathscr X^\infty$ for all possible $w\in\mathcal W(x)$.
%
The algorithm to determine the maximal robust positive invariant sets recursively intersects the current set iterate~$X_l$ with the set of all possible successor states starting from the current iterate $X_{l+1}=X_l\cap\left\{\bigcup_{x\in X_l,w\in\mathcal W(x)}\{Ax+w\}\right\}$.
%
We will only present a single step of this procedure.
%
Notice that for a constant set-valued map~$\mathcal W(x) = \mathcal W$ the iteration can be summarised as $X_{l+1} = X_l\cap A^{-1}(X_l\ominus\mathcal W)$, for a set valued map a similar compact notation also holds, however it is more confusing than helpful:~$X_{l+1}=X_l\cap(A^{-1}X_l\ominus A^{-1}\mathcal W(A^{-1}X_l))$.

@@

\begin{figure}
\centering
\begin{tikzpicture}[scale=1.8]
\draw[-latex'] (-2.1,0) -- (2.1,0) node[above] {$t$};
\draw[-latex'] (0,-.9) -- (0,.9) node[right,blue] {$f_1(t)$};
\foreach \x in {-2,-1.5556,...,2} \draw (\x,0.05) -- (\x,-.05);
\draw (-2,-.05) node[below] {$-2$};
\draw (2,-.05) node[below] {$2$};
\foreach \y in {-.8,-.4,...,.8} \draw (0.05,\y) -- (-.05,\y);
\draw (.05,-.8) node[right] {$-0.8$};
\draw (.05,-.4) node[right] {$-0.4$};
\draw[scale=1,domain=-2:2,smooth,variable=\x,blue] plot ({\x},{\x*\x*\x/10});
\draw (-2.0000, -0.8000) -- (-1.5556, -0.3764) -- (-1.1111, -0.1372) -- (-0.6667, -0.0296) -- (-0.2222, -0.0011) -- (0,0); 
\draw (0,0) -- (0.2222, 0.0011) -- (0.6667, 0.0296) -- (1.1111, 0.1372) -- (1.5556, 0.3764) -- (2.0000, 0.8000);
\fill[pattern = north east lines, opacity=.5] (-2.0000, -0.8000) -- (-1.5556, -0.3764) -- (-1.1111, -0.1372) -- (-0.6667, -0.0296) -- (-0.2222, -0.0011) -- (0,0) -- (-2,0) -- cycle; 
\fill[pattern = north east lines, opacity=.5] (0,0) -- (0.2222, 0.0011) -- (0.6667, 0.0296) -- (1.1111, 0.1372) -- (1.5556, 0.3764) -- (2.0000, 0.8000) -- (2,0) -- cycle;
\end{tikzpicture}
\caption{$f_1(t)$, for $t=\frac{x_1}{2}+x_2$, and its approximation by 9 secants used in Example~\ref{exp:second:example}.}
\label{fig:approximation:f1:p:diff}
\end{figure}

\begin{figure}
\centering
\begin{tikzpicture}[scale=1.8]
\draw[-latex'] (-2.1,0) -- (2.1,0) node[above] {$t$};
\draw[-latex'] (0,-.9) -- (0,.9) node[right,blue] {$f_2(t)$};
\foreach \x in {-2,-1.5556,...,2} \draw (\x,0.05) -- (\x,-.05);
\draw (-2,-.05) node[below] {$-2$};
\draw (2,-.05) node[below] {$2$};
\foreach \y in {-.8,-.4,...,.8} \draw (0.05,\y) -- (-.05,\y);
\draw (.05,-.8) node[right] {$-0.8$};
\draw (.05,-.4) node[right] {$-0.4$};
\draw[blue] (-2.0000, -0.7854) -- (-1.9596, -0.6847) -- (-1.9192, -0.6428) -- (-1.8788, -0.6104) -- (-1.8384, -0.5830) -- (-1.7980, -0.5587) -- (-1.7576, -0.5367) -- (-1.7172, -0.5163) -- (-1.6768, -0.4972) -- (-1.6364, -0.4791) -- (-1.5960, -0.4620) -- (-1.5556, -0.4456) -- (-1.5152, -0.4298) -- (-1.4747, -0.4146) -- (-1.4343, -0.3999) -- (-1.3939, -0.3856) -- (-1.3535, -0.3717) -- (-1.3131, -0.3581) -- (-1.2727, -0.3449) -- (-1.2323, -0.3319) -- (-1.1919, -0.3192) -- (-1.1515, -0.3068) -- (-1.1111, -0.2945) -- (-1.0707, -0.2825) -- (-1.0303, -0.2706) -- (-0.9899, -0.2589) -- (-0.9495, -0.2473) -- (-0.9091, -0.2359) -- (-0.8687, -0.2247) -- (-0.8283, -0.2135) -- (-0.7879, -0.2025) -- (-0.7475, -0.1915) -- (-0.7071, -0.1807) -- (-0.6667, -0.1699) -- (-0.6263, -0.1592) -- (-0.5859, -0.1486) -- (-0.5455, -0.1381) -- (-0.5051, -0.1276) -- (-0.4646, -0.1172) -- (-0.4242, -0.1069) -- (-0.3838, -0.0966) -- (-0.3434, -0.0863) -- (-0.3030, -0.0761) -- (-0.2626, -0.0658) -- (-0.2222, -0.0557) -- (-0.1818, -0.0455) -- (-0.1414, -0.0354) -- (-0.1010, -0.0253) -- (-0.0606, -0.0152) -- (-0.0202, -0.0051) -- ( 0.0202,  0.0503) -- ( 0.0606,  0.0875) -- ( 0.1010,  0.1133) -- ( 0.1414,  0.1346) -- ( 0.1818,  0.1531) -- ( 0.2222,  0.1699) -- ( 0.2626,  0.1854) -- ( 0.3030,  0.1999) -- ( 0.3434,  0.2136) -- ( 0.3838,  0.2267) -- ( 0.4242,  0.2393) -- ( 0.4646,  0.2515) -- ( 0.5051,  0.2633) -- ( 0.5455,  0.2747) -- ( 0.5859,  0.2859) -- ( 0.6263,  0.2969) -- ( 0.6667,  0.3077) -- ( 0.7071,  0.3184) -- ( 0.7475,  0.3289) -- ( 0.7879,  0.3393) -- ( 0.8283,  0.3496) -- ( 0.8687,  0.3598) -- ( 0.9091,  0.3699) -- ( 0.9495,  0.3801) -- ( 0.9899,  0.3902) -- ( 1.0303,  0.4003) -- ( 1.0707,  0.4104) -- ( 1.1111,  0.4205) -- ( 1.1515,  0.4307) -- ( 1.1919,  0.4410) -- ( 1.2323,  0.4513) -- ( 1.2727,  0.4618) -- ( 1.3131,  0.4723) -- ( 1.3535,  0.4830) -- ( 1.3939,  0.4939) -- ( 1.4343,  0.5050) -- ( 1.4747,  0.5164) -- ( 1.5152,  0.5280) -- ( 1.5556,  0.5400) -- ( 1.5960,  0.5523) -- ( 1.6364,  0.5651) -- ( 1.6768,  0.5785) -- ( 1.7172,  0.5926) -- ( 1.7576,  0.6076) -- ( 1.7980,  0.6237) -- ( 1.8384,  0.6413) -- ( 1.8788,  0.6610) -- ( 1.9192,  0.6842) -- ( 1.9596,  0.7141) -- ( 2.0000,  0.7854);
\draw (-2.0000, -0.7854) -- (-1.8000, -0.5599) -- (-1.5000, -0.4240) -- (0, 0);
\fill[pattern = north east lines, opacity=.5] (-2.0000, -0.7854) -- (-1.8000, -0.5599) -- (-1.5000, -0.4240) -- (0, 0) -- (-2,0) -- cycle;
\draw (-.5466,0) -- (0.7000, 0.3165) -- (1.0000, 0.3927) -- (1.5000, 0.5236) -- (1.7000, 0.5865) -- (2.0000, 0.7854);
\fill[pattern = north east lines, opacity=.5] (-.5466,0) -- (0.7000, 0.3165) -- (1.0000, 0.3927) -- (1.5000, 0.5236) -- (1.7000, 0.5865) -- (2.0000, 0.7854) -- (2,0) -- cycle;
\end{tikzpicture}
\caption{$f_2(t)$, for $t=x_1-\frac{x_2}{2}$, and its approximation by 5 secants and 4 tangents as used in Example~\ref{exp:second:example}.}
\label{fig:approximation:f2:p:diff}
\end{figure}
